An examination of the high-order dynamic interactions underlying multi-dimensional timeseries data

### 2018-10-16 paper/main.tex
added mexican hat kernel, working on levelup in methods [WIP]

\subsection*{Dynamic correlations}
Given a matrix of observations, we can compute the (static)
correlations between any pair of observations, $\mathbf{X}_i$ and
$\mathbf{X}_j$ using:
\begin{align}
  \mathrm{corr}(\mathbf{X}_i, \mathbf{X}_j) &= \frac{\sum_{t=1}^T \left(\mathbf{X}_i(t)
                                              -
                                              \hat{\mathbf{X}_i}\right) \left(\mathbf{X}_j(t)
                                              -
                                              \hat{\mathbf{X}_j}\right)}{\sqrt{\sum_{t=1}^T\left(
                                              \sigma^2_{\mathbf{X}_i}\right) \left(
                                              \sigma^2_{\mathbf{X}_j}\right)}},~\mathrm{where}\\
                                    \hat{\mathbf{X}_k} &= \sum_{t=1}^T
                                                         \mathbf{X}_k(t),~\mathrm{and}\\
  \sigma^2_{\mathbf{X}_k} &= \sum_{t=1}^T \left( \mathbf{X}_k -
                            \hat{\mathbf{X}_k} \right)^2 
\end{align}

We can generalize this formula to compute time-varying correlations by
incorporating a \textit{weight function} that takes a time $t$ as input,
and returns how much the observed data every timepoint (including $t$)
contribute to the correlations at time $t$ (Fig.~\ref{fig:weights}).
\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/kernels}
  \caption{\textbf{Examples of time-varying weights.  Each panel
    displays per-timepoint weights computed for 100 timepoints
    ($0, 1, ..., 99$) with $t = 50$.  a. Uniform weights.} The weights
  are timepoint-invariant; all timepoints are weighted equally, and do
  not change as a function of $t$.  This is a special case of weight
  function that reduces dynamic correlations to static correlations.
  \textbf{b. Dirac delta function.} Only the timepoint $t$ is given
  weight (of 1), and all other timepoints' weights are set to 0.
  \textbf{c. Gaussian weights.} Each timepoint's weights fall off
  according to a Gaussian probability density function centered on
  $\mu = t$.  Weights derived using several different example variance
  parameters ($\sigma^2$ are displayed.  \textbf{d. Laplace weights.}
  Each timepoint's weights fall off according to a Laplace probability
  density function centered on $\mu = t$.  Weights derived using
  several different example scale parameters ($b$) are displayed.}
  \label{fig:weights}
\end{figure}

%include figure showing weights (https://github.com/ContextLab/timecorr-paper/issues/1)

\textbf{JRM STOPPED HERE}
\begin{align}
\mathrm{corr}_t(x,y) &= \frac{\sum_{t=1}^T \left( x_t - \bar{x_t} \right) \left( y_t - \bar{y_t} \right)}{\sqrt{\sum_{t=1}^{T}\sigma ^{2}_{x_{t}}\sigma ^{2}_{y_{t}}}}\mathrm{, where}\\
\bar{a_t} &= \sum_{t=1}^T w(t)_i a_i\\
\sigma^{2}_{a_t} &= \sum_{t=1}^{T} \left(a_t - \bar{a_t} \right)^2\\
w(t) &= \mathcal{N}\left(1...T ~|~ t, \sigma_\mathcal{N}\right) %update to include eye, laplace, and gaussian weights
\end{align}

% \begin{equation}
% c(t) = \sum_{i=1}^{T}w(t)_{i}
% \end{equation}

>>>>

\subsection*{Dynamic correlations}
Given a matrix of observations, we can compute the (static)
correlations between any pair of observations, $\mathbf{X}_i$ and
$\mathbf{X}_j$ using:
\begin{align}
  \mathrm{corr}(\mathbf{X}_i, \mathbf{X}_j) &= \frac{\sum_{t=1}^T \left(\mathbf{X}_i(t)
                                              -
                                              \hat{\mathbf{X}_i}\right) \left(\mathbf{X}_j(t)
                                              -
                                              \hat{\mathbf{X}_j}\right)}{\sqrt{\sum_{t=1}^T
                                              \sigma^2_{\mathbf{X}_i} 
                                              \sigma^2_{\mathbf{X}_j}}},~\mathrm{where}\\\label{eqn:corr}
  \hat{\mathbf{X}_k} &= \sum_{t=1}^T
                       \mathbf{X}_k(t),~\mathrm{and}\\
  \sigma^2_{\mathbf{X}_k} &= \sum_{t=1}^T \left( \mathbf{X}_k -
                            \hat{\mathbf{X}_k} \right)^2 
\end{align}

We can generalize this formula to compute time-varying correlations by
incorporating a \textit{weight function} that takes a time $t$ as input,
and returns how much the observed data every timepoint (including $t$)
contribute to the correlations at time $t$ (Fig.~\ref{fig:weights}).
\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/kernels}
  \caption{\textbf{Examples of time-varying weights.  Each panel
      displays per-timepoint weights at $t = 50$, evaluated for 100
      timepoints ($1, ..., 100$).  a. Uniform weights.} The weights
    are timepoint-invariant; observations at all timepoints are
    weighted equally, and do not change as a function of $t$.  This is
    a special case of weight function that reduces dynamic
    correlations to static correlations.  \textbf{b. Dirac delta
      function.} Only the observation at timepoint $t$ is given weight
    (of 1), and weights for observations at all other timepoints are
    set to 0.  \textbf{c. Gaussian weights.} Each observation's
    weights fall off in time according to a Gaussian probability
    density function centered on $\mu = t$.  Weights derived using
    several different example variance parameters ($\sigma^2$ are
    displayed.  \textbf{d. Laplace weights.}  Each observation's
    weights fall off in time according to a Laplace probability
    density function centered on $\mu = t$.  Weights derived using
    several different example scale parameters ($b$) are displayed.
    \textbf{e. Mexican hat (Ricker wavelet) weights.}  Each
    observation's weights fall off in time according to a Ricker
    wavelet centered on $t$.  Weights derived using several different
    example width parameters ($\sigma$) are displayed.}
  \label{fig:weights}
\end{figure}

Given a weight function $w(t)$ for timepoint $t$, evaluated at
timepoints in the interval $\left[ 1, ..., T \right]$, we can extend the static correlation formula
in Equation~\ref{eqn:corr} to reflect an \textit{instantaneous
  correlation} at timepoint $t$:
\begin{align}
  \mathrm{timecorr}(\mathbf{X}_i, \mathbf{X}_j, t) &= \frac{\sum_{t=1}^T \left(\mathbf{X}_i(t)
                                              -
                                              \widetilde{\mathbf{X}}_i(t)\right) \left(\mathbf{X}_j(t)
                                              -
                                              \widetilde{\mathbf{X}}_j(t)\right)}{\sqrt{\sum_{t=1}^T
                                              \widetilde{\sigma}^2_{\mathbf{X}_i}(t) 
                                              \widetilde{\sigma}^2_{\mathbf{X}_j}}(t)},~\mathrm{where}\\\label{eqn:timecorr}
  \widetilde{\mathbf{X}}_k(t) &= \sum_{i=1}^T
                       w(t, i)\mathbf{X}_k(i),\\
  \widetilde{\sigma}^2_{\mathbf{X}_k}(t) &= \sum_{i=1}^T \left( \mathbf{X}_k(i) -
                            \widetilde{\mathbf{X}}_k(t) \right)^2,
\end{align}
and $w(t, i)$ is shorthand for $w(t)$ evaluated at timepoint $i$.
Equation~\ref{eqn:timecorr} may be used to estimate the instantaneous
correlations between every pair of observations, at each timepoint
(i.e., $\mathbf{Y}$).

\subsection*{Higher-order correlations}
Given a timeseries of dynamic correlations (e.g., obtained using
Eqn.~\ref{eqn:timecorr}), higher-order correlations reflect the
dynamic correlations between columns of $\mathbf{Y}$.  Given unlimited
computing resources, one could use repeated applications of
Equation~\ref{eqn:timecorr} to estimate these higher-order
correlations (i.e., substituting in the previous output, $\mathbf{Y}$,
for the input, $\mathbf{X}$ in the equation).  However, because each output
$\mathbf{Y}$ has $O(F^2)$ columns relative to $F$ columns in the input
$\mathbf{X}$, the output of Equation~\ref{eqn:timecorr} grows with the
square of the number of repeated applications.  When $F$ is large,
this approach quickly becomes intractable.

\textbf{JRM STOPPED HERE}

\subsection*{Timepoint decoding}

\subsection*{Prediction}
